---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
ggplot2::theme_set(ggplot2::theme_bw())
```

# Delphi Epidata R client


<!-- badges: start -->
[![License: MIT][mit-image]][mit-url] [![Github Actions][github-actions-image]][github-actions-url]
[![codecov](https://codecov.io/gh/dsweber2/epidatr/branch/dev/graph/badge.svg?token=jVHL9eHZNZ)](https://app.codecov.io/gh/dsweber2/epidatr)
<!-- badges: end -->


The [Delphi Epidatr package](https://cmu-delphi.github.io/epidatr/) is an R front-end for the [Delphi Epidata API](https://cmu-delphi.github.io/delphi-epidata/), which provides real-time access to epidemiological surveillance data for influenza, COVID-19, and other diseases for the USA at various geographical resolutions, both from official government sources such as the [Center for Disease Control (CDC)](https://www.cdc.gov/datastatistics/index.html) and [Google Trends](https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/google-symptoms.html) and private partners such as [Facebook](https://delphi.cmu.edu/blog/2020/08/26/covid-19-symptom-surveys-through-facebook/) and [Change Healthcare](https://www.changehealthcare.com/). It is built and maintained by the Carnegie Mellon University [Delphi research group](https://delphi.cmu.edu/).

This package is designed to streamline the downloading and usage of data from the [Delphi Epidata
API](https://cmu-delphi.github.io/delphi-epidata/). It provides a simple R interface to the API, including functions for downloading data, parsing the results, and converting the data into a tidy format. The API stores a historical record of all data, including corrections and updates, which is particularly useful for accurately backtesting forecasting models. We also provide packages for downstream data processing ([epiprocess](https://github.com/cmu-delphi/epiprocess)) and modeling ([epipredict](https://github.com/cmu-delphi/epipredict)).

## Usage
You can find detailed docs here: 

```{r}
library(epidatr)
# Obtain the smoothed covid-like illness (CLI) signal from the
# Facebook survey as it was on April 10, 2021 for the US
epidata <- pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_cli",
  geo_type = "nation",
  time_type = "day",
  geo_values = "us",
  time_values = epirange(20210101, 20210601),
  as_of = "2021-06-01"
)
epidata
```

```{r fb-cli-signal}
# Plot this data
library(ggplot2)
ggplot(epidata, aes(x = time_value, y = value)) +
  geom_line() +
  labs(
    title = "Smoothed CLI from Facebook Survey",
    subtitle = "US, 2021",
    x = "Date",
    y = "CLI"
  )
```


## Installation
You can install the stable version of this package from CRAN:

``` r
```R
install.packages("epidatr")
pak::pkg_install("epidatr")
renv::install("epidatr")
```
```

Or if you want the development version, install from GitHub:

```R
```R
# Install the dev version using `pak` or `remotes`
pak::pkg_install("cmu-delphi/epidatr")
remotes::install_github("cmu-delphi/epidatr")
```
```

### API Keys

The Delphi API requires a (free) API key for full functionality. To generate
your key, register for a pseudo-anonymous account
[here](https://api.delphi.cmu.edu/epidata/admin/registration_form) and see more
discussion on the [general API
website](https://cmu-delphi.github.io/delphi-epidata/api/api_keys.html). See the
`set_api_key()` function documentation for details on how to use your API key.

Note that the private endpoints (i.e. those prefixed with `pvt_`) require a
separate key that needs to be passed as an argument. These endpoints require
specific data use agreements to access.

[mit-image]: https://img.shields.io/badge/License-MIT-yellow.svg
[mit-url]: https://opensource.org/license/mit/
[github-actions-image]: https://github.com/cmu-delphi/epidatr/workflows/ci/badge.svg
[github-actions-url]: https://github.com/cmu-delphi/epidatr/actions

## Get updates

You should consider subscribing to the [API mailing list](https://lists.andrew.cmu.edu/mailman/listinfo/delphi-covidcast-api) to be notified of package updates, new data sources, corrections, and other updates.

## For users of the `covidcast` R package

The `epidatr` package is a complete rewrite of the [`covidcast` package](https://cmu-delphi.github.io/covidcast/covidcastR/), with a focus on speed, reliability, and ease of use. The `covidcast` package is deprecated and will no longer be updated.
