---
title: "Delphi Epidata R API Client"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Delphi Epidata R API Client}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{ggplot2}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L, max.print = 4L)
library(epidatr)
library(dplyr)
```

The epidatr package provides access to all the endpoints of the [Delphi Epidata
API](https://cmu-delphi.github.io/delphi-epidata/), and can be used to make
requests for specific signals on specific dates and in selected geographic
regions.

We recommend you register for an API key. While most endpoints are available
without one, there are [limits on API usage for anonymous
users](https://cmu-delphi.github.io/delphi-epidata/api/api_keys.html), including
a rate limit. If you regularly request large amounts of data, please consider
[registering for an API
key](https://api.delphi.cmu.edu/epidata/admin/registration_form). You can then
specify this key by either

- setting the environment variable `DELPHI_EPIDATA_KEY`, such as by editing your
  `.Renviron` file
- setting the option `delphi.epidata.key` using `options()`

## Basic Usage

Fetching some data from the Delphi Epidata API is simple. Suppose we are
interested in the [`covidcast`
endpoint](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html), which
provides access to a range of data on COVID-19. Reviewing the endpoint
documentation, we see that we need to specify a data source name, a signal name,
a geographic level, a time resolution, and the location and times of interest.

In this case, the `pub_covidcast()` function lets us specify these parameters for
the endpoint and returns a tibble with the results:

```{r}
epidata <- pub_covidcast(
  "fb-survey", "smoothed_cli", "state", "day", "pa",
  epirange(20210105, 20210410)
)
epidata
```

We can then easily plot the data using ggplot2:

```{r, out.height="65%"}
library(ggplot2)
ggplot(epidata, aes(x = time_value, y = value)) +
  geom_line() +
  labs(
    title = "Smoothed CLI from Facebook Survey",
    subtitle = "PA, 2021",
    x = "Date",
    y = "CLI"
  )
```

The [Delphi Epidata API documentation](https://cmu-delphi.github.io/delphi-epidata/) has more information on the available endpoints and arguments. You can also use the `avail_endpoints()` function to get a table of endpoint functions:

```{r}
avail_endpoints()
```

Example queries with all the endpoint functions available in this package are given [below](#example-queries).

## Advanced Usage (Experimental)

The [COVIDcast
endpoint](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html) of the
Epidata API contains many separate data sources and signals. It can be difficult
to find the name of the signal you're looking for, so you can use
`covidcast_epidata()` to get help with finding sources and functions without
leaving R.

The `covidcast_epidata()` function fetches a list of all signals, and returns an
object containing fields for every signal:

```{r}
epidata <- covidcast_epidata()
epidata$signals
```

If you use an editor that supports tab completion, such as RStudio, type
`epidata$signals$` and wait for the tab completion popup. You will be able to
type the name of signals and have the autocomplete feature select them from the
list for you. Note that some signal names have dashes in them, so to access them
we rely on the backtick operator:

```{r}
epidata$signals$`fb-survey:smoothed_cli`
```

These objects can be used directly to fetch data, without requiring us to use
the `pub_covidcast()` function. Simply use the `$call` attribute of the object:

```{r}
epidata$signals$`fb-survey:smoothed_cli`$call("state", "pa", epirange(20210405, 20210410))
```

## Advanced Usage (Debugging)

We can obtain the [`epidata_call`] object underlying a request by setting the
`dry_run` argument to `TRUE` in `fetch_args_list()`:

```{r}
pub_covidcast(
  "fb-survey", "smoothed_cli", "state", "day", "pa",
  epirange(20210405, 20210410),
  fetch_args = fetch_args_list(dry_run = TRUE)
)
```

## Example Queries

(Some endpoints allow for the use of `*` to access data at all locations. Check the help for a given endpoint to see if it supports `*`.)

### COVIDcast Main Endpoint

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html>

County geo_values are [FIPS codes](https://en.wikipedia.org/wiki/List_of_United_States_FIPS_codes_by_county) and are discussed in the API docs [here](https://cmu-delphi.github.io/delphi-epidata/api/covidcast_geography.html). The example below is for Orange County, California.

```{r}
pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_accept_covid_vaccine",
  geo_type = "county",
  time_type = "day",
  time_values = epirange(20201221, 20201225),
  geo_values = "06059"
)
```

The `covidcast` endpoint supports `*` in its time and geo fields:

```{r}
pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_accept_covid_vaccine",
  geo_type = "county",
  time_type = "day",
  time_values = epirange(20201221, 20201225),
  geo_values = "*"
)
```

### Other Covid Endpoints

#### COVID-19 Hospitalization: Facility Lookup

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covid_hosp_facility_lookup.html>

```{r, eval = FALSE}
pub_covid_hosp_facility_lookup(city = "southlake")
pub_covid_hosp_facility_lookup(state = "WY")
# A non-example (there is no city called New York in Wyoming)
pub_covid_hosp_facility_lookup(state = "WY", city = "New York")
```

#### COVID-19 Hospitalization by Facility

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covid_hosp_facility.html>

```{r, eval = FALSE}
pub_covid_hosp_facility(
  hospital_pks = "100075",
  collection_weeks = epirange(20200101, 20200501)
)
```

#### COVID-19 Hospitalization by State

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covid_hosp.html>

```{r, eval = FALSE}
pub_covid_hosp_state_timeseries(states = "MA", dates = "20200510")
```

### Flu Endpoints

#### Delphi's ILINet forecasts

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/delphi.html>

```{r, eval = FALSE}
del <- pub_delphi(system = "ec", epiweek = 201501)
names(del[[1L]]$forecast)
```

#### FluSurv hospitalization data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/flusurv.html>

```{r, eval = FALSE}
pub_flusurv(locations = "ca", epiweeks = 202001)
```

#### Fluview data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/fluview.html>

```{r, eval = FALSE}
pub_fluview(regions = "nat", epiweeks = epirange(201201, 202001))
```

#### Fluview virological data from clinical labs

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/fluview_clinical.html>

```{r, eval = FALSE}
pub_fluview_clinical(regions = "nat", epiweeks = epirange(201601, 201701))
```

#### Fluview metadata

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/fluview_meta.html>

```{r, eval = FALSE}
pub_fluview_meta()
```

#### Google Flu Trends data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/gft.html>

```{r, eval = FALSE}
pub_gft(locations = "hhs1", epiweeks = epirange(201201, 202001))
```

#### ECDC ILI

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/ecdc_ili.html>

```{r, eval = FALSE}
pub_ecdc_ili(regions = "Armenia", epiweeks = 201840)
```

#### KCDC ILI

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/kcdc_ili.html>

```{r, eval = FALSE}
pub_kcdc_ili(regions = "ROK", epiweeks = 200436)
```

#### NIDSS Flu

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/nidss_flu.html>

```{r, eval = FALSE}
pub_nidss_flu(regions = "taipei", epiweeks = epirange(200901, 201301))
```

#### ILI Nearby Nowcast

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/nowcast.html>

```{r, eval = FALSE}
pub_nowcast(locations = "ca", epiweeks = epirange(202201, 202319))
```

### Dengue Endpoints

#### Delphi's Dengue Nowcast

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/dengue_nowcast.html>

```{r, eval = FALSE}
pub_dengue_nowcast(locations = "pr", epiweeks = epirange(201401, 202301))
```

#### NIDSS dengue

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/nidss_dengue.html>

```{r, eval = FALSE}
pub_nidss_dengue(locations = "taipei", epiweeks = epirange(200301, 201301))
```

### PAHO Dengue

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/paho_dengue.html>

```{r, eval=FALSE}
pub_paho_dengue(regions = "ca", epiweeks = epirange(200201, 202319))
```

### Other Endpoints

#### Wikipedia Access

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/wiki.html>

```{r, eval = FALSE}
pub_wiki(language = "en", articles = "influenza", epiweeks = epirange(202001, 202319))
```

### Private methods

These require private access keys to use (separate from the Delphi Epidata API key).
To actually run these locally, you will need to store these secrets in your `.Reviron` file, or set them as environmental variables.

#### CDC

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/cdc.html>

```{r, eval=FALSE}
pvt_cdc(auth = Sys.getenv("SECRET_API_AUTH_CDC"), epiweeks = epirange(202003, 202304), locations = "ma")
```

#### Dengue Digital Surveillance Sensors

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/dengue_sensors.html>

```{r, eval=FALSE}
pvt_dengue_sensors(
  auth = Sys.getenv("SECRET_API_AUTH_SENSORS"),
  names = "ght",
  locations = "ag",
  epiweeks = epirange(201404, 202004)
)
```

#### Google Health Trends

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/ght.html>

```{r, eval=FALSE}
pvt_ght(
  auth = Sys.getenv("SECRET_API_AUTH_GHT"),
  epiweeks = epirange(199301, 202304),
  locations = "ma",
  query = "how to get over the flu"
)
```

#### NoroSTAT metadata

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/meta_norostat.html>

```{r, eval=FALSE}
pvt_meta_norostat(auth = Sys.getenv("SECRET_API_AUTH_NOROSTAT"))
```

#### NoroSTAT data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/norostat.html>

```{r, eval=FALSE}
pvt_norostat(auth = Sys.getenv("SECRET_API_AUTH_NOROSTAT"), locations = "1", epiweeks = 201233)
```

#### Quidel Influenza testing

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/quidel.html>

```{r, eval=FALSE}
pvt_quidel(auth = Sys.getenv("SECRET_API_AUTH_QUIDEL"), locations = "hhs1", epiweeks = epirange(200301, 202105))
```

#### Sensors

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/sensors.html>

```{r, eval=FALSE}
pvt_sensors(
  auth = Sys.getenv("SECRET_API_AUTH_SENSORS"),
  names = "sar3",
  locations = "nat",
  epiweeks = epirange(200301, 202105)
)
```

#### Twitter

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/twitter.html>

```{r, eval=FALSE}
pvt_twitter(
  auth = Sys.getenv("SECRET_API_AUTH_TWITTER"),
  locations = "nat",
  epiweeks = epirange(200301, 202105)
)
```
