---
title: "Get started with epidatr"
output:
  rmarkdown::html_vignette:
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Get started with epidatr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{ggplot2}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L, max.print = 4L)
```

The epidatr package provides access to all the endpoints of the [Delphi Epidata
API](https://cmu-delphi.github.io/delphi-epidata/), and can be used to make
requests for specific signals on specific dates and in select geographic
regions.


## Setup

### Installation

You can install the stable version of this package from CRAN:

```R
install.packages("epidatr")
pak::pkg_install("epidatr")
renv::install("epidatr")
```

Or if you want the development version, install from GitHub:

```R
# Install the dev version using `pak` or `remotes`
pak::pkg_install("cmu-delphi/epidatr@dev")
remotes::install_github("cmu-delphi/epidatr", ref = "dev")
renv::install("cmu-delphi/epidatr@dev")
```

### API Keys

The Delphi API requires a (free) API key for full functionality. While most
endpoints are available without one, there are
[limits on API usage for anonymous users](https://cmu-delphi.github.io/delphi-epidata/api/api_keys.html),
including a rate limit.

To generate your key,
[register for a pseudo-anonymous account](https://api.delphi.cmu.edu/epidata/admin/registration_form).
See the `save_api_key()` function documentation for details on how to set up
`epidatr` to use your API key.

_Note_ that private endpoints (i.e. those prefixed with `pvt_`) require a
separate key that needs to be passed as an argument. These endpoints require
specific data use agreements to access.


## Basic Usage

Fetching data from the Delphi Epidata API is simple. Suppose we are
interested in the
[`covidcast` endpoint](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html),
which provides access to a
[wide range of data](https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html)
on COVID-19. Reviewing the endpoint documentation, we see that we
[need to specify](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html#constructing-api-queries)
a data source name, a signal name, a geographic level, a time resolution, and
the location and times of interest.

The `pub_covidcast()` function lets us access the `covidcast` endpoint:

```{r}
library(epidatr)
library(dplyr)

# Obtain the most up-to-date version of the smoothed covid-like illness (CLI)
# signal from the COVID-19 Trends and Impact survey for the US
epidata <- pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_cli",
  geo_type = "nation",
  time_type = "day",
  geo_values = "us",
  time_values = epirange(20210105, 20210410)
)
knitr::kable(head(epidata))
```

`pub_covidcast()` returns a `tibble`. (Here we’re using `knitr::kable()` to make
it more readable.) Each row represents one observation in Pennsylvania on one
day. The state abbreviation is given in the `geo_value` column, the date in
the `time_value` column. Here `value` is the requested signal -- in this
case, the smoothed estimate of the percentage of people with COVID-like
illness, based on the symptom surveys, and `stderr` is its standard error.

The Epidata API makes signals available at different geographic levels,
depending on the endpoint. To request signals for all states instead of the
entire US, we use the `geo_type` argument paired with `*` for the
`geo_values` argument. (Only some endpoints allow for the use of `*` to
access data at all locations. Check the help for a given endpoint to see if
it supports `*`.)

```{r, eval = FALSE}
# Obtain the most up-to-date version of the smoothed covid-like illness (CLI)
# signal from the COVID-19 Trends and Impact survey for all states
pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_cli",
  geo_type = "state",
  time_type = "day",
  geo_values = "*",
  time_values = epirange(20210105, 20210410)
)
```

We can fetch a subset of states by listing out the desired locations:

```{r, eval = FALSE}
# Obtain the most up-to-date version of the smoothed covid-like illness (CLI)
# signal from the COVID-19 Trends and Impact survey for Pennsylvania
pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_cli",
  geo_type = "state",
  time_type = "day",
  geo_values = c("pa", "ca", "fl"),
  time_values = epirange(20210105, 20210410)
)
```

We can also request data for a single location at a time, via the `geo_values` argument.

```{r}
# Obtain the most up-to-date version of the smoothed covid-like illness (CLI)
# signal from the COVID-19 Trends and Impact survey for Pennsylvania
epidata <- pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_cli",
  geo_type = "state",
  time_type = "day",
  geo_values = "pa",
  time_values = epirange(20210105, 20210410)
)
knitr::kable(head(epidata))
```

## Getting versioned data

The Epidata API stores a historical record of all data, including corrections
and updates, which is particularly useful for accurately backtesting
forecasting models. To fetch versioned data, we can use the `as_of`
argument.

```{r, eval = FALSE}
# Obtain the smoothed covid-like illness (CLI) signal from the COVID-19
# Trends and Impact survey for Pennsylvania as it was on 2021-06-01
pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_cli",
  geo_type = "state",
  time_type = "day",
  geo_values = "pa",
  time_values = epirange(20210105, 20210410),
  as_of = "2021-06-01"
)
```

See `vignette("versioned-data")` for details and more ways to specify versioned data.


## Plotting

Because the output data is in a standard `tibble` format, we can easily plot
it using `ggplot2`:

```{r, out.height="65%"}
library(ggplot2)
ggplot(epidata, aes(x = time_value, y = value)) +
  geom_line() +
  labs(
    title = "Smoothed CLI from Facebook Survey",
    subtitle = "PA, 2021",
    x = "Date",
    y = "CLI"
  )
```

`ggplot2` can also be used to [create choropleths](https://r-graphics.org/recipe-miscgraph-choropleth).


```{r, class.source = "fold-hide", out.height="65%"}
library(maps)

# Obtain the most up-to-date version of the smoothed covid-like illness (CLI)
# signal from the COVID-19 Trends and Impact survey for all states on a single day
cli_states <- pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_cli",
  geo_type = "state",
  time_type = "day",
  geo_values = "*",
  time_values = 20210410
)

# Get a mapping of states to longitude/latitude coordinates
states_map <- map_data("state")

# Convert state abbreviations into state names
cli_states <- mutate(
  cli_states,
  state = ifelse(
    geo_value == "dc",
    "district of columbia",
    state.name[match(geo_value, tolower(state.abb))] %>% tolower()
  )
)

# Add coordinates for each state
cli_states <- left_join(states_map, cli_states, by = c("region" = "state"))

# Plot
ggplot(cli_states, aes(x = long, y = lat, group = group, fill = value)) +
  geom_polygon(colour = "black", linewidth = 0.2) +
  coord_map("polyconic") +
  labs(
    title = "Smoothed CLI from Facebook Survey",
    subtitle = "All states, 2021-04-10",
    x = "Longitude",
    y = "Latitude"
  )
```


## Finding locations of interest

Most data is only available for the US. Select endpoints report other countries at the national and/or regional levels. Endpoint descriptions explicitly state when they cover non-US locations.

For endpoints that report US data, see the
[geographic coding documentation](https://cmu-delphi.github.io/delphi-epidata/api/covidcast_geography.html)
for available geographic levels.


### International data

International data is available via

- `pub_dengue_nowcast` (North and South America)
- `pub_ecdc_ili` (Europe)
- `pub_kcdc_ili` (Korea)
- `pub_nidss_dengue` (Taiwan)
- `pub_nidss_flu` (Taiwan)
- `pub_paho_dengue` (North and South America)
- `pvt_dengue_sensors` (North and South America)


## Finding data sources and signals of interest

Above we used data from [Delphi’s symptom surveys](https://delphi.cmu.edu/covid19/ctis/),
but the Epidata API includes numerous data streams: medical claims data, cases
and deaths, mobility, and many others. This can make it a challenge to find
the data stream that you are most interested in.

The Epidata documentation lists all the data sources and signals available
through the API for [COVID-19](https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html)
and for [other diseases](https://cmu-delphi.github.io/delphi-epidata/api/README.html#source-specific-parameters).

You can also use the `avail_endpoints()` function to get a table of endpoint functions:

```{r, eval = FALSE}
avail_endpoints()
```

```{r, echo = FALSE}
invisible(capture.output(endpts <- avail_endpoints()))
knitr::kable(endpts)
```

See `vignette("signal-discovery")` for more information.
