---
title: "Migrating from pub_covidcast to the new Epidata API"
output:
  rmarkdown::html_vignette:
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Migrating from pub_covidcast to the new Epidata API}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = identical(Sys.getenv("NOT_CRAN"), "true"))
```

```{r setup, message = FALSE, eval = TRUE}
library(epidatr)
```

The Delphi Epidata API is moving from the covidcast endpoint (API v4, served
by `pub_covidcast()`) to a new set of endpoints (API v5, served by
`epidata_snapshot()`, `epidata_archive()`, and `epidata_meta()`). The
transition is in progress: sources are moving to the new API one at a time,
and `pub_covidcast()` still works for sources that have not moved yet. New
analyses should start with the new functions and fall back to
`pub_covidcast()` only when a source is not yet available there.

This guide maps the old interface onto the new one.

## Function mapping

| Old | New | Purpose |
|---|---|---|
| `pub_covidcast()` | `epidata_snapshot()` | Data as it appeared on a given date (or the latest) |
| `pub_covidcast(issues = ...)` | `epidata_archive()` | Full revision history of a signal |
| `pub_covidcast_meta()`, `covidcast_epidata()` | `epidata_meta()` | Discover sources, signals, geo types, and date ranges |

`epidata()` is a convenience wrapper that routes to `epidata_snapshot()` or
`epidata_archive()` based on which versioning argument you pass.

## Argument changes

| `pub_covidcast()` argument | New argument | Notes |
|---|---|---|
| `source`, `signals`, `geo_type`, `geo_values` | same | |
| `time_type` | none | Dropped. Times in the new API are always `Date`s. |
| `time_values` | `reference_time` | Accepts dates or `epirange()`. Filtered locally after the fetch. |
| `as_of` | `snapshot_date` | `epidata_snapshot()` only. `NULL` returns the latest data. |
| `issues` | `report_time` | `epidata_archive()` only. Accepts exact dates, operators like `"<2025-10-16"`, or `epirange()`. |
| `lag` | none | Compute it yourself: `report_time - reference_time`. |

The new functions also add `fill_method`, which has no covidcast equivalent.
Some sources publish several variants of the same signal that differ in how
nulls were handled during geographic aggregation: `"source"` (raw source
data, no imputation), `"fill_ave"` (nulls filled with the average of
neighboring values), and `"fill_zero"` (nulls filled with zero). The default
`NULL` returns all variants, so filter on this column (or pass the argument)
if you want exactly one time series per location.

## Column changes

| `pub_covidcast()` column | New column | Notes |
|---|---|---|
| `geo_value`, `geo_type`, `signal`, `value` | same | |
| `time_value` | `reference_time` | The date the value describes. Always a `Date`. |
| `issue` | `report_time` | The date the value was published. Present in both snapshot and archive output. |
| `source` | dropped | You queried by source; add it back with `dplyr::mutate()` if you bind rows across sources. |
| `time_type` | dropped | No longer needed since times are `Date`s. |
| `lag` | dropped | Compute as `report_time - reference_time`. |
| `direction` | dropped | Was already deprecated in the covidcast API. |
| `stderr`, `sample_size` | `ci_lower`, `ci_upper` | Uncertainty is now expressed as confidence interval bounds on `value` instead of a standard error. Populated only for sources that publish them. See below. |
| `missing_value`, `missing_stderr`, `missing_sample_size` | dropped | Missingness is now expressed through `fill_method` variants and plain `NA`s. |
| none | `fill_method` | Which null-handling variant of the signal this row belongs to. See above. |

Some sources also carry extra columns in the new API, for example `age_group`
(pophive) and `nwss_source`, `sample_index`, `pcr_target` (nwss).

### Uncertainty columns

The covidcast columns `stderr` and `sample_size` have no fixed replacement.
The shared schema carries only `value`; a source that quantifies uncertainty
adds its own columns, such as `ci_lower` and `ci_upper`. Use the metadata or 
the [documentation](https://cmu-delphi.github.io/delphi-epidata/api/v5_signals.html)
 to see which value columns a source returns:

```{r value-columns}
meta_sleepcycle <- epidata_meta(source = "sleepcycle")
meta_sleepcycle$sleepcycle$value_columns
```

## A query, before and after

Fetching NSSP influenza ED visit percentages for two states, as the data
looked on January 1, 2025:

```{r before}
old <- pub_covidcast(
  source = "nssp",
  signals = "pct_ed_visits_influenza",
  geo_type = "state",
  time_type = "week",
  geo_values = c("pa", "ca"),
  time_values = epirange(202440, 202501),
  as_of = 20250101
)
head(old)
```

```{r after}
new <- epidata_snapshot(
  source = "nssp",
  signals = "pct_ed_visits_influenza",
  geo_type = "state",
  geo_values = c("pa", "ca"),
  reference_time = epirange("2024-10-01", "2025-01-01"),
  snapshot_date = "2025-01-01"
)
head(new)
```

## Revision history queries

Where you used to pass `issues` to `pub_covidcast()`, use
`epidata_archive()` with `report_time`:

```{r archive}
revisions <- epidata_archive(
  source = "nssp",
  signals = "pct_ed_visits_influenza",
  geo_type = "state",
  geo_values = "pa",
  report_time = "<2025-06-01"
)
head(revisions)
```

If you filtered by `lag`, fetch the archive and filter afterwards:

```{r lag, eval = FALSE}
revisions[revisions$report_time - revisions$reference_time <= 7, ]
```

## Checking whether a source has moved

Use `epidata_meta()` to see what a source offers in the new API. It returns
signals, geo types, and the available `reference_time` and `report_time`
ranges:

```{r meta}
meta <- epidata_meta(source = "nssp")
meta$nssp$signals
meta$nssp$time_value_range
```

If `epidata_meta()` does not know the source yet, keep using
`pub_covidcast()` for it and check back after package updates. The
[API mailing list](https://lists.andrew.cmu.edu/mailman/listinfo/delphi-covidcast-api)
announces sources as they move.
