---
title: "Finding data sources and signals of interest"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Finding data sources and signals of interest}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L, max.print = 4L)
library(epidatr)
library(dplyr)
```

The Epidata API includes numerous data streams -- medical claims data, cases and deaths, mobility, and many others -- covering different geographic regions. This can make it a challenge to find the data stream that you are most interested in.

Example queries with all the endpoint functions available in this package are
given [below](#example-queries).


## Using the documentation

The Epidata documentation lists all the data sources and signals available
through the API for
[COVID-19](https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html) and
for [other diseases](https://cmu-delphi.github.io/delphi-epidata/api/README.html#source-specific-parameters).
The site also includes a search tool if you have a keyword (e.g. "Taiwan") in mind.


## Signal metadata

Some endpoints have partner metadata available that provides information about
the signals that are available, for example, what time ranges they are
available for, and when they have been updated.

```{r, echo = FALSE}
suppressMessages(invisible(capture.output(endpts <- avail_endpoints())))
filter(endpts, endsWith(Endpoint, "_meta()")) %>% knitr::kable()
```

## Interactive tooling

We provide a couple `epidatr` functions to help find data sources and signals.

The `avail_endpoints()` function lists endpoints, each of which, except for
COVIDcast, corresponds to a single data source. `avail_endpoints()` outputs a
`tibble` of endpoints and brief descriptions, which explicitly state when they
cover non-US locations:

```{r, eval = FALSE}
avail_endpoints()
```

```{r, echo = FALSE}
suppressMessages(invisible(capture.output(endpts <- avail_endpoints())))
knitr::kable(endpts)
```

The `covidcast_epidata()` function lets you look more in-depth at the data
sources available through the COVIDcast endpoint. The function describes
all available data sources and signals:

```{r}
covid_sources <- covidcast_epidata()
head(covid_sources$sources, n = 2)
```

Each source is included as an entry in the `covid_sources$sources` list, associated
with a `tibble` describing included signals.

If you use an editor that supports tab completion, such as RStudio, type
`covid_sources$source$` and wait for the tab completion popup. You will be able to
browse the list of data sources.

```{r}
covid_sources$signals
```

If you use an editor that supports tab completion, type
`covid_sources$signals$` and wait for the tab completion popup. You will be
able to type the name of signals and have the autocomplete feature select
them from the list for you. In the tab-completion popup, signal names are
prefixed with the name of the data source for filtering convenience.

_Note_ that some signal names have dashes in them, so to access them
we rely on the backtick operator:

```{r}
covid_sources$signals$`fb-survey:smoothed_cli`
```

These signal objects can be used directly to fetch data, without requiring us to use
the `pub_covidcast()` function. Simply use the `$call` attribute of the object:

```{r}
epidata <- covid_sources$signals$`fb-survey:smoothed_cli`$call(
  "state", "pa", epirange(20210405, 20210410)
)
knitr::kable(epidata)
```


## Example Queries

### COVIDcast Main Endpoint

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html>

County geo_values are [FIPS codes](https://en.wikipedia.org/wiki/List_of_United_States_FIPS_codes_by_county) and are discussed in the API docs [here](https://cmu-delphi.github.io/delphi-epidata/api/covidcast_geography.html). The example below is for Orange County, California.

```{r}
pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_accept_covid_vaccine",
  geo_type = "county",
  time_type = "day",
  time_values = epirange(20201221, 20201225),
  geo_values = "06059"
)
```

The `covidcast` endpoint supports `*` in its time and geo fields:

```{r}
pub_covidcast(
  source = "fb-survey",
  signals = "smoothed_accept_covid_vaccine",
  geo_type = "county",
  time_type = "day",
  time_values = epirange(20201221, 20201225),
  geo_values = "*"
)
```

### Other Covid Endpoints

#### COVID-19 Hospitalization: Facility Lookup

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covid_hosp_facility_lookup.html>

```{r, eval = FALSE}
pub_covid_hosp_facility_lookup(city = "southlake")
pub_covid_hosp_facility_lookup(state = "WY")
# A non-example (there is no city called New York in Wyoming)
pub_covid_hosp_facility_lookup(state = "WY", city = "New York")
```

#### COVID-19 Hospitalization by Facility

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covid_hosp_facility.html>

```{r, eval = FALSE}
pub_covid_hosp_facility(
  hospital_pks = "100075",
  collection_weeks = epirange(20200101, 20200501)
)
```

#### COVID-19 Hospitalization by State

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/covid_hosp.html>

```{r, eval = FALSE}
pub_covid_hosp_state_timeseries(states = "MA", dates = "20200510")
```

### Flu Endpoints

#### Delphi's ILINet forecasts

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/delphi.html>

```{r, eval = FALSE}
del <- pub_delphi(system = "ec", epiweek = 201501)
names(del[[1L]]$forecast)
```

#### FluSurv hospitalization data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/flusurv.html>

```{r, eval = FALSE}
pub_flusurv(locations = "ca", epiweeks = 202001)
```

#### Fluview data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/fluview.html>

```{r, eval = FALSE}
pub_fluview(regions = "nat", epiweeks = epirange(201201, 202001))
```

#### Fluview virological data from clinical labs

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/fluview_clinical.html>

```{r, eval = FALSE}
pub_fluview_clinical(regions = "nat", epiweeks = epirange(201601, 201701))
```

#### Fluview metadata

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/fluview_meta.html>

```{r, eval = FALSE}
pub_fluview_meta()
```

#### Google Flu Trends data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/gft.html>

```{r, eval = FALSE}
pub_gft(locations = "hhs1", epiweeks = epirange(201201, 202001))
```

#### ECDC ILI

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/ecdc_ili.html>

```{r, eval = FALSE}
pub_ecdc_ili(regions = "Armenia", epiweeks = 201840)
```

#### KCDC ILI

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/kcdc_ili.html>

```{r, eval = FALSE}
pub_kcdc_ili(regions = "ROK", epiweeks = 200436)
```

#### NIDSS Flu

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/nidss_flu.html>

```{r, eval = FALSE}
pub_nidss_flu(regions = "taipei", epiweeks = epirange(200901, 201301))
```

#### ILI Nearby Nowcast

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/nowcast.html>

```{r, eval = FALSE}
pub_nowcast(locations = "ca", epiweeks = epirange(202201, 202319))
```

### Dengue Endpoints

#### Delphi's Dengue Nowcast

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/dengue_nowcast.html>

```{r, eval = FALSE}
pub_dengue_nowcast(locations = "pr", epiweeks = epirange(201401, 202301))
```

#### NIDSS dengue

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/nidss_dengue.html>

```{r, eval = FALSE}
pub_nidss_dengue(locations = "taipei", epiweeks = epirange(200301, 201301))
```

### PAHO Dengue

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/paho_dengue.html>

```{r, eval=FALSE}
pub_paho_dengue(regions = "ca", epiweeks = epirange(200201, 202319))
```

### Other Endpoints

#### Wikipedia Access

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/wiki.html>

```{r, eval = FALSE}
pub_wiki(
  language = "en",
  articles = "influenza",
  time_type = "week",
  time_values = epirange(202001, 202319)
)
```

### Private methods

These require private access keys to use (separate from the Delphi Epidata API key).
To actually run these locally, you will need to store these secrets in your `.Reviron` file, or set them as environmental variables.

<details id="private-endpoints">

<summary>Usage of private endpoints</summary>

#### CDC

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/cdc.html>

```{r, eval=FALSE}
pvt_cdc(auth = Sys.getenv("SECRET_API_AUTH_CDC"), epiweeks = epirange(202003, 202304), locations = "ma")
```

#### Dengue Digital Surveillance Sensors

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/dengue_sensors.html>

```{r, eval=FALSE}
pvt_dengue_sensors(
  auth = Sys.getenv("SECRET_API_AUTH_SENSORS"),
  names = "ght",
  locations = "ag",
  epiweeks = epirange(201404, 202004)
)
```

#### Google Health Trends

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/ght.html>

```{r, eval=FALSE}
pvt_ght(
  auth = Sys.getenv("SECRET_API_AUTH_GHT"),
  epiweeks = epirange(199301, 202304),
  locations = "ma",
  query = "how to get over the flu"
)
```

#### NoroSTAT metadata

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/meta_norostat.html>

```{r, eval=FALSE}
pvt_meta_norostat(auth = Sys.getenv("SECRET_API_AUTH_NOROSTAT"))
```

#### NoroSTAT data

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/norostat.html>

```{r, eval=FALSE}
pvt_norostat(auth = Sys.getenv("SECRET_API_AUTH_NOROSTAT"), locations = "1", epiweeks = 201233)
```

#### Quidel Influenza testing

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/quidel.html>

```{r, eval=FALSE}
pvt_quidel(auth = Sys.getenv("SECRET_API_AUTH_QUIDEL"), locations = "hhs1", epiweeks = epirange(200301, 202105))
```

#### Sensors

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/sensors.html>

```{r, eval=FALSE}
pvt_sensors(
  auth = Sys.getenv("SECRET_API_AUTH_SENSORS"),
  names = "sar3",
  locations = "nat",
  epiweeks = epirange(200301, 202105)
)
```

#### Twitter

API docs: <https://cmu-delphi.github.io/delphi-epidata/api/twitter.html>

```{r, eval=FALSE}
pvt_twitter(
  auth = Sys.getenv("SECRET_API_AUTH_TWITTER"),
  locations = "nat",
  time_type = "week",
  time_values = epirange(200301, 202105)
)
```

</details>
