---
title: "epidatr v5 API demo"
output:
  rmarkdown::html_vignette:
    code_folding: show
vignette: >
  %\VignetteIndexEntry{epidatr v5 API demo}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, message = FALSE, eval = TRUE}
library(dplyr)
library(ggplot2)
library(epidatr)
```

### Metadata

Let's check the source-specific metadata for `nssp`.

```{r meta-example}
meta_nssp <- epidata_meta(source = "nssp")
meta_nssp$nssp$signals
meta_nssp$nssp$geo_types
meta_nssp$nssp$version_range
meta_nssp$nssp$time_value_range
```

### Basic Queries

We can pull the latest snapshot of a signal.

```{r snapshot-example}
nssp_data <- epidata_snapshot(
  source = "nssp",
  signal = "pct_ed_visits_influenza",
  geo_type = "state"
)
head(nssp_data)
```

If you want to inspect the API request URL or query structure without actually fetching the data, you can use the `dry_run` argument via `fetch_args_list()`:

```{r dry-run-example}
dry_run_call <- epidata_snapshot(
  source = "nssp",
  signal = "pct_ed_visits_influenza",
  geo_type = "state",
  fetch_args = fetch_args_list(dry_run = TRUE)
)
dry_run_call
```

Filtering by specific geographies and versions:

```{r geo-filter}
pa_ca_data <- epidata_snapshot(
  source = "nssp",
  signal = "pct_ed_visits_influenza",
  geo_type = "state",
  geo_values = c("PA", "CA"),
  as_of = "2025-01-01" # fetch data as it was known on this date
)
head(pa_ca_data)
```

### Archive Queries

If you want to track how data for a specific time period was revised over time, you can use `epidata_archive()`.

```{r archive-example}
archive_data <- epidata_archive(
  source = "nssp",
  signal = "pct_ed_visits_influenza",
  geo_type = "state"
)
head(archive_data)
```

### Other Sources

Here are some examples for NHSN (hospitalizations), POPHIVE, and NWSS (wastewater).

```{r other-sources-examples}
# NHSN: Hospital Admissions
meta_nhsn <- epidata_meta(source = "nhsn")
meta_nhsn$nhsn$signals
meta_nhsn$nhsn$geo_types
meta_nhsn$nhsn$version_range
meta_nhsn$nhsn$time_value_range
nhsn_data <- epidata_snapshot(
  source = "nhsn",
  signal = "confirmed_admissions_flu_ew",
  geo_type = "state"
)
head(nhsn_data)

# POPHIVE
meta_pophive <- epidata_meta(source = "pophive")
meta_pophive$pophive$signals
meta_pophive$pophive$geo_types
meta_pophive$pophive$version_range
meta_pophive$pophive$time_value_range
pophive_data <- epidata_snapshot(
  source = "pophive",
  signal = "covid_pct_ed",
  geo_type = "state"
)
head(pophive_data)

# NWSS: Wastewater Surveillance
meta_nwss <- epidata_meta(source = "nwss")
meta_nwss$nwss$signals
meta_nwss$nwss$geo_types
meta_nwss$nwss$version_range
meta_nwss$nwss$time_value_range
nwss_data <- epidata_snapshot(
  source = "nwss",
  signal = "covid_avg_conc",
  geo_type = "sewershed"
)
head(nwss_data)
```
